Visualizations for the manuscript

### 2022-03-12 metanalysis_visualization/manuscript_Figures.Rmd
figures updated for the manuscript

```{r}
# cut-off dates used in the analysis, these should not be changed
time_period <-"months" #possible values: days, weeks, months
history_start_date <- as.Date("2019-01-01")
history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-01-15")
end_date_plots <- as.Date("2021-05-01")
```

@@

patients_hospitalized_agg_sex_country <- patients_hospitalized_agg_sex_perc_all %>%
  select( - percentage ) %>%
  inner_join( country_map ) %>%
  group_by( time_p, sex, country ) %>%
  mutate( total_cases = sum( count ), 
          sex_cases = sum( count_sex ), 
          percentage = 100* sex_cases / total_cases) %>%
  rbind( patients_hospitalized_agg_sex_total)

patients_hospitalized_agg_sex_country %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
  geom_line( aes(linetype = sex, color = sex )) +
  geom_point( aes( shape= sex, color = sex )) +
  scale_colour_manual(values=c("#FF6600","#339999")) +
  scale_fill_manual(values=c("#FF6600","#339999"))+
  #scale_y_continuous(breaks = seq(0, 30, by = 1)) +
  facet_wrap2("country", axes = "all") + 
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA), 
        legend.position="bottom")
```

@@

perc_disorder_group_total %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
  geom_line( aes(color = disorder_group )) +
  geom_point( aes( color = disorder_group )) +
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))

perc_disorder_group_country <- perc_disorder_group_all %>%
  select( - percentage_dg ) %>%
  inner_join( country_map ) %>%
  group_by( time_p, disorder_group, country ) %>%
  mutate( total_cases = sum( count_icd ), 
          dg_cases = sum( count_dg ), 
          percentage = 100* dg_cases / total_cases ) %>%
  rbind( perc_disorder_group_total)

perc_disorder_group_country %>%
  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
  "Mental Health Symptoms", "Obsessive-Compulsive and Related Disorders" )) %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
  geom_line( aes(linetype = disorder_group, color = disorder_group )) +
  geom_point( aes( shape= disorder_group, color = disorder_group )) +
  scale_y_continuous(breaks = seq(0, 30, by = 1)) +
  facet_wrap2("country", axes = "all") + 
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA), 
        legend.position="bottom")

perc_disorder_group_country %>%
  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
  "Mental Health Symptoms", "Obsessive-Compulsive and Related Disorders" )) %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
  geom_line( aes(linetype = country, color = country )) +
  geom_point( aes( shape= country, color = country )) +
  scale_y_continuous(breaks = seq(0, 30, by = 5)) +
  facet_wrap2("disorder_group", axes = "all") + 
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
  scale_color_manual( values = cbPalette ) + 
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA), 
        legend.position="bottom")

@@

>>>>

```{r}
# cut-off dates used in the analysis, these should not be changed
time_period <-"months" #possible values: days, weeks, months
history_start_date <- as.Date("2019-01-01")
history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-01-01")
end_date_plots <- as.Date("2021-05-31")
```

@@

#patients_hospitalized_agg_sex_country <- patients_hospitalized_agg_sex_perc_all %>%
#  select( - percentage ) %>%
#  inner_join( country_map ) %>%
#  group_by( time_p, sex, country ) %>%
#  mutate( total_cases = sum( count ), 
#          sex_cases = sum( count_sex ), 
#          percentage = 100* sex_cases / total_cases) %>%
#  rbind( patients_hospitalized_agg_sex_total)

# patients_hospitalized_agg_sex_country %>%
#   ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
#   geom_line( aes(linetype = sex, color = sex )) +
#   geom_point( aes( shape= sex, color = sex )) +
#   scale_colour_manual(values=c("#FF6600","#339999")) +
#   scale_fill_manual(values=c("#FF6600","#339999"))+
#   #scale_y_continuous(breaks = seq(0, 30, by = 1)) +
#   facet_wrap2("country", axes = "all") + 
#   geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
#   labs(x ="", y = "percentage of patients (%)")+
#   scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
#   theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
#         strip.text = element_text(size = 10, face = "bold"),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
#         panel.border = element_rect(colour = "black", fill = NA), 
#         legend.position="bottom")
```

@@

perc_disorder_group_total %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
  geom_line( aes(color = disorder_group )) +
  geom_point( aes( color = disorder_group )) +
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))
#   
# perc_disorder_group_country <- perc_disorder_group_all %>%
#   select( - percentage_dg ) %>%
#   inner_join( country_map ) %>%
#   group_by( time_p, disorder_group, country ) %>%
#   mutate( total_cases = sum( count_icd ), 
#           dg_cases = sum( count_dg ), 
#           percentage = 100* dg_cases / total_cases ) %>%
#   rbind( perc_disorder_group_total)

perc_disorder_group_total %>%
  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
  "Mental Health Symptoms", "Feeding and Eating Disorders" )) %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
  geom_line( aes(linetype = country, color = country )) +
  geom_point( aes( shape= country, color = country )) +
  #scale_y_continuous(breaks = seq(0, 30, by = 5)) +
  facet_wrap2("disorder_group", axes = "all", scales ="free") + 
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
  scale_color_manual( values = cbPalette ) + 
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA), 
        legend.position="bottom")

@@

## ITS with seasonality

### By site
```{r}
head(patient_count_psy_period_psy_all)
sites <- unique( patient_count_psy_period_psy_all$siteid )
source("../metanalysis_visualization/time_series_analysis_plotting.R")
p <- list()

for( i in 1:length(sites)){
  print(i)
  subset_site <- patient_count_psy_period_psy_all %>%
    filter( siteid == sites[i])
  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
                              subset_site,
                              "percentage",
                              clearance_period = FALSE) +
  labs(title = sites[i],
       x = "Calendar dates",
       y = "Percentage") +
     theme(strip.text = element_text(size = 10),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
        plot.subtitle = element_text(size = 12, face = "italic"))

}

g <- do.call(grid.arrange,p)
ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
